# Supplemental Material

```{r, echo=F}
trans_pal = c("Ag Expansion: Cropland" = "Peru",
              "Ag Expansion: Pasture" = "LightYellow",
              "Urbanization: High" = "Black",
              "Urbanization: Medium" = "DimGray",
              "Urbanization: Low" = "DarkGray",
              "Urbanization: Open" = "LightGray",
              "Forest Harvest: Forest Clearcut" = "Purple4",
              "Forest Harvest: Forest Selection" = "Purple1",
              "Fire: High Severity" = "DarkRed",
              "Fire: Medium Severity" = "Red",
              "Fire: Low Severity" = "Salmon",
              "Insect: High Severity" = "DarkGreen",
              "Insect: Medium Severity" = "Green3",
              "Insect: Low Severity" = "LightGreen") 

```

```{r, diagurb, cache=T, echo=F, warning=F, fig.width=3, Fig.height=3, fig.cap="Conceptual diagram of urbanization and urban intensification transitions used within the LUCAS state-and-transition simulation model."}

DiagrammeR::grViz("digraph rmarkdown{
graph [layout = dot, rankdir = LR]

node [shape = rectangle, fontname = 'Roboto Condensed', fixedsize = true, width=1]

node [style = filled; fillcolor = white]
Forest; Grassland; Shrubland; Wetland; Cropland; Pasture

node [shape = rectangle; style = filled; fillcolor = Gray90]
Urbanization

node [shape = rectangle; style = filled; fillcolor = Red4]
High

node [shape = rectangle; style = filled; fillcolor = OrangeRed]
Medium

node [shape = rectangle; style = filled; fillcolor = LightSalmon]
Low

node [shape = rectangle; style = filled; fillcolor = MistyRose]
Open

edge [arrowhead = vee, color = Red]
{Forest; Grassland; Shrubland; Wetland; Cropland; Pasture} -> Urbanization
Urbanization -> {High; Medium; Low; Open}

edge [arrowhead = normal, color = Black]
{Open} -> {Low; Medium; High}
{Low} -> {Medium; High}
{Medium} -> {High}

subgraph {rank = same; Open; Low; Medium; High}

}")
```

```{r, diagharv, cache=T, echo=F, warning=F, fig.width=3, Fig.height=3, fig.cap="Conceptual diagram of forest harvest transitions used within the LUCAS state-and-transition simulation model."}

DiagrammeR::grViz("digraph rmarkdown{
graph [layout = dot, rankdir = TB]

node [fontname = 'Roboto Condensed', fixedsize = true, fontsize = 8]
edge [fontname= 'Roboto Condensed', fontsize = 7]

node [shape = rectangle, style = filled; fillcolor = ForestGreen]
Forest

node [shape = rectangle; style = filled; fillcolor = Gray90]
Clearcut; Selection

edge [arrowhead = vee, color = black, font = 'Universe Condensed']
Forest -> {Harvest} [label = 'Annual area']
Harvest -> Clearcut [label = 'Harvest multipliers', style = dashed]
Harvest -> Selection [style = dashed]

Clearcut -> Forest [label = 'regrowth']
Selection -> Forest

subgraph {rank = same; Clearcut; Selection}

}")
```

```{r, diagfire, cache=T, echo=F, warning=F, fig.width=3, Fig.height=3, fig.cap="Conceptual diagram of wildfire transitions used within the LUCAS state-and-transition simulation model."}

DiagrammeR::grViz("digraph rmarkdown{
graph [layout = dot, rankdir = TB]

node [fontname = 'Roboto Condensed', fixedsize = true, fontsize = 8]
edge [fontname= 'Roboto Condensed', fontsize = 7]

node [shape = rectangle, style = filled; fillcolor = ForestGreen]
Forest

node [shape = rectangle, style = filled; fillcolor = '#DCCA8F']
Shrubland


node [shape = rectangle; style = filled; fillcolor = Gray90]
Fire

node [shape = rectangle; style = filled; fillcolor = Gray90]
High; Medium; Low

edge [arrowhead = vee, color = black, font = 'Universe Condensed']
Forest -> Fire [label = 'Annual Perimeters']
Fire -> High [style = dashed]
Fire -> Medium [label = 'severity multiplier'; style = dashed]
Fire -> Low [style = dashed]

{Medium; Low} -> Forest
High -> Shrubland
Shrubland -> Forest [label = 'regrowth multiplier'; style = dashed]

Forest -> Selection
Selection -> Prescribed
Prescribed -> Low
Prescribed -> Medium

subgraph {rank = same; High; Medium; Low}
subgraph {rank = same; Shrubland; Forest}

}")
```


```{r, stockresults, cache=T, echo=F, warning=F, message=F, fig.width=8, fig.height=4, fig.cap="Carbon stocks in conterminous U.S. forests from 2001-2020."}

ggplot(stocks_ipcc, aes(x=Timestep, y=Amount/1000000, color=ScenarioName, fill=ScenarioName, shape=ScenarioName)) +
  geom_line() +
  scale_color_brewer(palette="Set2") +
  geom_point(color="white", size=1.5) +
  scale_fill_brewer(palette="Set2") +
  scale_shape_manual(values=c("LULC+Clim"=21, "Clim"=22, "LULC"=23, "Ref"=24)) +
  facet_wrap(~StockGroupID, scales="free_y", ncol=5) +
  theme_custom() +
  labs(x="Year", y=expression(Carbon~storage~(MMT~C)), fill="Scenario", shape="Scenario", color="Scenario") +
    theme(legend.position = "bottom",
        legend.key.width = unit(10, "mm"),
        legend.key.height = unit(3, "mm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,-10,0,-10))
```

```{r, stocksstateclass, echo=F, warning=F, message=F, fig.width=6, fig.height=6, fig.cap="Net change in carbon stock for each forest type and state class."}

stocks_state_class = read_csv("data/results/stocks-state-class.csv") %>%
  left_join(scenario_ids) %>%
  filter(StockGroupID %in% c("Biomass: Aboveground", "Biomass: Belowground", "DOM: Deadwood", "DOM: Litter", "DOM: Soil"), Timestep %in% c(2001,2020), ScenarioName=="LULC+Clim") %>%
  group_by(ScenarioName, StockGroupID, StateClassID) %>%
  arrange(Timestep) %>%
  mutate(Change = Amount-lag(Amount)) %>%
  filter(Timestep==2020, Change!=0) %>%
  mutate(StateClassID = str_remove(StateClassID, " Group"))

tec_state_class = stocks_state_class %>%
  group_by(StateClassID) %>% summarise(Amount=sum(Amount), Change=sum(Change)) %>%
  mutate(StockGroupID = "Total Ecosystem")

ggplot() +
  geom_bar(data=stocks_state_class, aes(x=StateClassID, y=Change/1000000, fill=StockGroupID), stat="identity") +
  geom_segment(data=tec_state_class, aes(x=StateClassID, xend=StateClassID, yend=0, y=Change/1000000, color="TEC")) +
  geom_point(data=tec_state_class, aes(x=StateClassID, y=Change/1000000, color="TEC"), shape=16) +
  scale_color_manual(values = c("TEC"="black")) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = comma) +
  coord_flip() +
  labs(x=NULL, y=expression(MMT~C), color=NULL, fill="Stock Group") +
  theme_custom() +
  theme(legend.position = c(0.8,0.1),
        legend.direction = "vertical",
        legend.key.height = unit(2, "mm"),
        legend.margin = margin(-1,-1,-10,-10))

```

Total ecosystem carbon (TEC) is the sum of all carbon stored in ecosystem pools (i.e. above and belowground biomass, deadwood, litter, and soil). The areas with the largest carbon stores were generally in the western mountain regions which often stored in excess of 300 tons per hectare.

```{r, tecmap, cache=T, echo=F, warning=F, message=F, fig.width=8, fig.cap="Map of the total ecosystem carbon stored in conterminous U.S. forests in the year 2020."}
land_poly = st_read("data/vector/ne_10m_land.shp", quiet=T)
states_poly = st_read("data/vector/states.shp", quiet=T)
tec_raster = raster("data/results/stocks-tec-2020.tif")
# tec_raster = aggregate(tec_raster, fact=4, na.rm=T)
tec_raster_df = as.data.frame(tec_raster, xy=T) %>%
  rename("Amount"="stocks.tec.2020") %>%
  filter(Amount>0)

breaks = c(min(tec_raster_df$Amount),50,100,150,200,300,max(tec_raster_df$Amount))
tec_raster_df = tec_raster_df %>% mutate(Class = cut(Amount, breaks=breaks, right=F))
unique(tec_raster_df$Class)

# min(tec_raster_df$Amount)
# max(tec_raster_df$Amount)
# quantile(tec_raster_df$Amount, probs = seq(0, 1, length.out = 9 + 1))

ggplot() +
  geom_sf(data=land_poly, aes(geometry=geometry), fill="white", color=NA) +
  geom_tile(data=tec_raster_df, aes(x=x, y=y, fill=Amount)) +
  geom_sf(data=states_poly, aes(geometry=geometry), fill=NA, color="grey20", size=0.2) +
  coord_sf(crs = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs",
           xlim=c(min(tec_raster_df$x), max(tec_raster_df$x)),
           ylim=c(min(tec_raster_df$y), max(tec_raster_df$y))) +
  scale_fill_fermenter(palette = "YlGnBu", type="seq", direction=1, guide="colorsteps", breaks=c(50,100,130,160,190,250,300,400)) + 
  #scale_fill_manual(values=c("grey90", "grey50", "lightyellow", "orange", "forestgreen", "purple"), guide="colorsteps") +
  labs(x=NULL, y=NULL, fill=expression(Metric~Tons~C~ha^-1)) +
  theme_bw() +
  theme(text=element_text(family="Roboto Condensed"),
        legend.position = c(0.2,0.1),
        legend.text.align = 0,
        legend.text = element_text(size = 7, hjust = 0, color = "#4e4d47"),
        legend.title = element_text(size = 8),
        plot.margin = unit(c(1,1,1,1), "mm"),
        legend.key.height = unit(2, "mm"),
        legend.key.width = unit(10, "mm"),
        axis.text.x = element_text(size=8),
        axis.text.y = element_text(size=8),
        panel.background = element_rect(fill="grey90")) +
  guides(fill = guide_colorsteps(direction = "horizontal",
                                 title.position = 'top',
                                 title.hjust = 0,
                                 label.hjust = 1,
                                 nrow = 1,
                                 byrow = T,
                                 reverse = F,
                                 label.position = "bottom",
                                 ticks = T,
                                 ticks.colour = "black",
                                 ticks.linewidth = 0.8,
                                 frame.colour = "black"))

```

```{r, stateclasstecchange, cache=T, echo=F, warning=F, message=F, fig.width=8, fig.height=4}
stocks_state_class = read_csv("data/results/stocks-state-class.csv") %>%
  left_join(scenario_ids) %>%
  filter(ScenarioName == "LULC+Clim", StockGroupID == "Total Ecosystem", str_detect(StateClassID, "Forest:")) %>%
  mutate(Amount=Amount/1000000) %>%
  filter(Timestep %in% c(2001,2020)) %>%
  group_by(StateClassID) %>%
  arrange(Timestep) %>%
  mutate(Change = Amount-lag(Amount)) %>%
  filter(Timestep == 2020)
stocks_state_class

ggplot(stocks_state_class, aes(x=fct_reorder(StateClassID, Change), y=Change, fill=StateClassID)) +
  geom_bar(stat="identity", show.legend = F) +
  coord_flip() +
  theme_custom() +
  labs(x=NULL, y=expression(Net~flux~(MMT~C)), fill="Scenario", shape="Scenario", color="Scenario") +
  theme(legend.position = c(0.85,0.2),
        legend.key.width = unit(5, "mm"),
        legend.key.height = unit(3, "mm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,-10,0,-10))

```



```{r}

###########
# Maps of NBP
###########
states_raster = raster("data/raster/ic-states.tif")
filter = states_raster %in% c(4,6,8,16,30,32,35,41,49,53,56)
western_states_raster = mask(states_raster, filter, maskvalue=0)

projection_raster = stack("data/results/canesm-flows-nbp-2050.tif", "data/results/hadgem2-flows-nbp-2050.tif")
projection_raster = mask(projection_raster, western_states_raster)
plot(projection_raster)

projection_raster_df = as.data.frame(projection_raster, xy=T) %>%
  rename("CanESM"="canesm.flows.nbp.2050", "HadGEM2"="hadgem2.flows.nbp.2050") %>%
  filter(!is.na(CanESM), !is.na(HadGEM2)) %>%
  pivot_longer(cols=c(-x, -y), names_to="GCM", values_to="Amount")

land_poly = st_read("data/vector/ne_10m_land.shp", quiet=T)
states_poly = st_read("data/vector/cb_2018_us_state_20m.shp", quiet=T) %>%
  filter(STUSPS %in% c("CA","OR","WA","ID","MT","WY","NV","UT","CO","NM","AZ"))

p2 = ggplot() +
  geom_sf(data=land_poly, aes(geometry=geometry), fill="white", color=NA) +
  geom_tile(data=projection_raster_df, aes(x=x, y=y, fill=Amount)) +
  geom_sf(data=states_poly, aes(geometry=geometry), fill=NA, color="grey20", size=0.2) +
  coord_sf(crs = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs",
           xlim=c(min(projection_raster_df$x), max(projection_raster_df$x)),
           ylim=c(min(projection_raster_df$y), max(projection_raster_df$y))) +
  scale_fill_fermenter(palette = "BrBG", type="seq", direction=1, guide="colorsteps", breaks=c(-100,-20,-10,-5,-2,2,5,10,20, 50)) +
  facet_wrap(~GCM) +
  labs(x=NULL, y=NULL, fill=expression(Metric~Tons~C~ha^-1)) +
  theme_bw(10) +
  theme(#text=element_text(family="Roboto Condensed"),
    legend.position = "bottom",
    legend.text.align = 0,
    legend.text = element_text(size = 7, hjust = 0, color = "#4e4d47"),
    legend.title = element_text(size = 8),
    plot.margin = unit(c(1,1,1,1), "mm"),
    legend.key.height = unit(2, "mm"),
    legend.key.width = unit(8, "mm"),
    legend.background = element_blank(),
    axis.text.x = element_text(size=8),
    axis.text.y = element_text(size=8),
    panel.background = element_rect(fill="grey90")) +
  guides(fill = guide_colorsteps(direction = "horizontal",
                                 title.position = 'top',
                                 title.hjust = 0,
                                 label.hjust = 1,
                                 nrow = 1,
                                 byrow = T,
                                 reverse = F,
                                 label.position = "bottom",
                                 ticks = T,
                                 ticks.colour = "black",
                                 ticks.linewidth = 0.8,
                                 frame.colour = "black"))

p = plot_grid(p2,p1, ncol = 2, align=c("hv"), rel_widths = c(6,4))
p
```


## Regional variability

The largest net carbon losses were in found in the mountain regions of California and Oregon (Sierra Nevada, Klamath and Cascades ecoregions) and were largely driven by stand-replacing wildfires and forest harvest. While remaining a net sink of carbon, the southeast experienced declines due to high rates of forest harvest and urbanization. However, high rates of productivity in the region, along with reforestation of agricultural lands, offset carbon losses ultimately resulting in a modest net sink. Texas was the state with the largest net sink, largely the result of forest aging and low rates of stand-replacing disturbances over the study period.

```{r, nbpmap, cache=T, echo=F, warning=F, message=F, fig.width=8, fig.cap="Map of the estiamted net biome productivity of conterminous U.S. forests for the period 2001-2020. Negative values indicate a net loss of carbon from ecosystems and positive values indicate a net sink of carbon."}
land_poly = st_read("data/vector/ne_10m_land.shp", quiet=T)
states_poly = st_read("data/vector/cb_2018_us_state_20m.shp", quiet=T)
nbp_raster = raster("data/results/flows-nbp-2020.tif")
# nbp_raster = aggregate(nbp_raster, fact=4, na.rm=T)
nbp_raster_df = as.data.frame(nbp_raster, xy=T) %>%
  rename("Amount"="flows.nbp.2020") %>%
  filter(!is.na(Amount))
# 
# min(nbp_raster_df$Amount)
# max(nbp_raster_df$Amount)
# quantile(nbp_raster_df$Amount, probs = seq(0, 1, length.out = 9 + 1))

p1 = ggplot() +
  geom_sf(data=land_poly, aes(geometry=geometry), fill="white", color=NA) +
  geom_tile(data=nbp_raster_df, aes(x=x, y=y, fill=Amount)) +
  geom_sf(data=states_poly, aes(geometry=geometry), fill=NA, color="grey20", size=0.2) +
  coord_sf(crs = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs",
           xlim=c(min(nbp_raster_df$x), max(nbp_raster_df$x)),
           ylim=c(min(nbp_raster_df$y), max(nbp_raster_df$y))) +
  scale_fill_fermenter(palette = "BrBG", type="seq", direction=1, guide="colorsteps", breaks=c(-100,-20,-10,-5,-2,2,5,10,20, 50)) +
  labs(x=NULL, y=NULL, fill=expression(Metric~Tons~C~ha^-1)) +
  theme_bw() +
  theme(#text=element_text(family="Roboto Condensed"),
        legend.position = c(0.2,0.1),
        legend.text.align = 0,
        legend.text = element_text(size = 7, hjust = 0, color = "#4e4d47"),
        legend.title = element_text(size = 8),
        plot.margin = unit(c(1,1,1,1), "mm"),
        legend.key.height = unit(2, "mm"),
        legend.key.width = unit(8, "mm"),
        legend.background = element_blank(),
        axis.text.x = element_text(size=8),
        axis.text.y = element_text(size=8),
        panel.background = element_rect(fill="grey90")) +
  guides(fill = guide_colorsteps(direction = "horizontal",
                                 title.position = 'top',
                                 title.hjust = 0,
                                 label.hjust = 1,
                                 nrow = 1,
                                 byrow = T,
                                 reverse = F,
                                 label.position = "bottom",
                                 ticks = T,
                                 ticks.colour = "black",
                                 ticks.linewidth = 0.8,
                                 frame.colour = "black"))

```

```{r, stratumflows, cache=T, echo=F, warning=F, message=F, fig.width=8, fig.height=3, fig.cap="Mean annual net biome productivity (NBP) of conterminous U.S. forests for the period 2002-2020 summarized by ecoregion (a) and states (b). Panel c shows NBP estimated for all forest cells inthe conterminous U.S."}
# Ecoregion chloropleth map NBP ----

eco_poly = st_read("data/vector/us_eco_l3_no_st_simp.shp", quiet=T) %>%
  st_simplify()
eco_code = read_csv("data/tables/ecoregions.csv")

flows_ecoreg = read_csv("data/results/flows_ecoregions.csv") %>%
  left_join(scenario_ids) %>%
  mutate(Amount = Amount/1000000) %>%
  filter(FlowGroupID == "Net Biome Productivity", ScenarioName == "LULC+Clim") %>%
  group_by(ScenarioName, FromStratumID, FlowGroupID) %>%
  summarise(Mean=mean(Amount), Min=quantile(Amount, 0.025), Max=quantile(Amount, 0.975)) %>%
  arrange(Mean) %>%
  left_join(eco_code, by=c("FromStratumID"="Name")) %>%
  mutate(ID = as.character(ID)) %>%
  left_join(eco_poly, by=c("ID"="US_L3CODE"))

p2 = ggplot(flows_ecoreg, aes(geometry=geometry, fill = Mean)) +
  geom_sf(color="black", size=0.1) +
  coord_sf(crs = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs") +
  #scale_fill_fermenter(palette = "BrBG", direction=1 , type="seq", guide="colorsteps", breaks=c(-4,-2,-1,-0.5,0.5,1,2,4,8)) +
  scale_fill_gradient2(low=muted("red"), mid="white", high=muted("green"), midpoint=0, guide="colorsteps", n.breaks=9) +
  theme_bw(9) +
  labs(fill=expression(MMT~C~yr^-1)) +
  theme(legend.key.height = unit(5, "mm"),
        legend.key.width = unit(2, "mm"),
        plot.margin = margin(2,2,2,2, "mm")) +
  guides(fill = guide_colorsteps(ticks = T,
                                 ticks.colour = "black",
                                 ticks.linewidth = 0.8,
                                 frame.colour = "black"))

# State chloropleth map NBP ----

state_poly = st_read("data/vector/cb_2018_us_state_20m.shp", quiet=T) %>% 
  mutate(STATEFP = as.character(STATEFP)) %>%
  st_transform(crs=crs)
state_code = read_csv("data/tables/states.csv")

flows_states = read_csv("data/results/flows_states.csv") %>%
  left_join(scenario_ids) %>%
  mutate(Amount = Amount/1000000) %>%
  filter(FlowGroupID == "Net Biome Productivity", ScenarioName == "LULC+Clim") %>%
  group_by(ScenarioName, FromSecondaryStratumID, FlowGroupID) %>%
  summarise(Mean=mean(Amount), Min=quantile(Amount, 0.025), Max=quantile(Amount, 0.975)) %>%
  arrange(Mean) %>%
  left_join(state_code, by=c("FromSecondaryStratumID"="Name")) %>%
  mutate(ID = as.character(ID)) %>%
  left_join(state_poly, by=c("ID"="STATEFP"))

p3 = ggplot(flows_states, aes(geometry=geometry, fill=Mean)) +
  geom_sf(color="black", size=0.1) +
  coord_sf(crs = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs") +
  #scale_fill_fermenter(palette = "BrBG", direction=1 , type="seq", guide="colorsteps", breaks=c(-4,-2,-1,-0.5,0.5,1,2,4,8)) +
  scale_fill_gradient2(low=muted("red"), mid="white", high=muted("green"), midpoint=0, guide="colorsteps", n.breaks=9) +
  theme_bw(9) +
  labs(fill=expression(MMT~C~yr^-1)) +
  theme(legend.key.height = unit(5, "mm"),
        legend.key.width = unit(2, "mm"),
        plot.margin = margin(2,2,2,2, "mm")) +
  guides(fill = guide_colorsteps(ticks = T,
                                 ticks.colour = "black",
                                 ticks.linewidth = 0.8,
                                 frame.colour = "black"))

# Create the plot ----
pleft = plot_grid(p2,p3, align=c("hv"), ncol = 1, labels=c("a", "b"), label_size = 10, label_fontface = "plain")
pright = plot_grid(p1, align=c("hv"), ncol = 1, labels=c("c"), label_size = 10, label_fontface = "plain")
p = plot_grid(pleft, pright, align=c("hv"), ncol = 2, rel_widths = c(4,6))
p
```
