--- 
title: "Spatially explicit estimates of carbon stocks and fluxes for the United States: a new approach linking the LUCAS and CBM-CFS3 models"
author: "Benjamin M. Sleeter"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib, references.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

Running Title: Ecosystem carbon balance in US Forests

Benjamin M. Sleeter^1^*, Leonardo Frid^2^, Bronwyrn Rayfield^2^, Paul C. Selmants^3^, Jinxun Liu^3^, Colin J. Daniel^2^

**Author Affiliations** 
^1^U.S. Geological Survey, Seattle, WA, USA; bsleeter@usgs.gov; (253) 343-3363   
^2^Apex Resource Management Solutions Ltd., Ottawa, ON, CAN  
^3^U.S. Geological Survey, Menlo Park, CA, USA

*Corresponding author

# Abstract

**Background**

**Results**

**Conclusions**

**Keywords**

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

```{r}
library(DT)
library(tidyverse)
library(scales)
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.name/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
