---
title: "Spatially explicit estimates of carbon stocks and fluxes for the United States: a new approach linking the LUCAS and CBM-CFS3 models"
author: "Benjamin M. Sleeter"
date: "`r Sys.Date()`"
bibliography: references.bib
csl: apa.csl
output:
  md_document: github_document
  html_notebook: default
  word_document: default
  pdf_document: default
  html_document:
    theme: readable 
    df_print: paged
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(DT)
library(kableExtra)
```

Running Title: Ecosystem carbon balance in US Forests

Benjamin M. Sleeter^1^*, Leonardo Frid^2^, Bronwyrn Rayfield^2^, Paul C. Selmants^3^, Jinxun Liu^3^, Colin J. Daniel^2^

#### Author Affiliations 
^1^U.S. Geological Survey, Seattle, WA, USA; bsleeter@usgs.gov; (253) 343-3363   
^2^Apex Resource Management Solutions Ltd., Ottawa, ON, CAN  
^3^U.S. Geological Survey, Menlo Park, CA, USA   

## Abstract

**Background**

**Results**

**Conclusions**

**Keywords**


## Background

P1: Paragraph discussing the importance and need for ecosystem carbon balance estimates and projections. 
•	What are the important considerations (e.g. LULC change, disturbances, climate variability and change)?
•	Why are these estimates needed (e.g. meeting local to global ghg-reduction targets).
•	What are the existing limitations of current approaches other than the CBM and LUCAS models?
P2: Review of the CBM-CFS2 approach and its major limitations.
P3: Review of the LUCAS approach and its major limitations.
P4: Describe in high-level terms what the aim of this study was.


This research describes a new approach to estimating land use and land cover change and their effect on ecosystem carbon dynamics by linking the CBM-CFS3 and LUCAS models to produce spatially explicit projections of ecosystem carbon balance in forested ecosystems of the conterminous United States.

The LUCAS model (Land Use and Carbon Scenario Simulator) is a spatially explicit state-and-transition simulation model (Daniel et al, 2016; Daniel et al., 2018) developed to assess the impact of land use and land cover change, ecosystem disturbance, and climate change on carbon stocks and fluxes. LUCAS has been used to explore and impacts of a range of plausible future scenarios on land use and carbon stocks (Sleeter et al., 2018, Sleeter et al., 2019). LUCAS runs within the SyncroSim software modeling environment (http://syncrosim.com).

The CBM-CFS3 model (Carbon Budget Model of the Canadian Forest Sector v3) is an IPCC Tier 3 spatially referenced model of carbon stocks and fluxes which can be applied at multiple scales (Kurz et al., 2009). The CBM-CFS3 model estimates the impacts of a range of land-use and disturbances on carbon stocks and fluxes. The approach relies on the conversion of merchantable volume curves (user supplied) into carbon biomass through the use of expansion factors. Turnover of live carbon and decay and decomposition of dead organic matter (DOM) are specified for each forest type and/or at a regional scale. CBM-CFS3 has been applied and adapted in many other countries using regionally-specific parameters. It's transparent approach and robust methodology make it an idea model to adapt for the U.S. forest sector.

## Results

## Discussion

## Conlucsion

## Methods

Model Structure
To estimate carbon stocks and fluxes for CONUS, we linked the CBM-CFS3 spatially referenced model of ecosystem carbon dynamics with the LUCAS spatially explicit model of LULC change. The CBM-CFS3 model was used to generate a set of forest species and ecoregion-level carbon stock and flux rates which were then used within the LUCAS model to produce spatially explicit maps of carbon stocks and fluxes based on LULC change, disturbances, forest ageing, and climate variability.

The conterminous U.S (CONUS) was partitioned into a regular grid of 1-km x 1-km cells where each cell was assigned to a discrete land-use/land-cover classes (LULC). The initial land cover map was based on the 2001 National Land Cover Database (NLCD). We further modified the classification system to partition the three NLCD forest classes into forest type-groups based on the U.S. Forest Service classification system. Furthermore, areas classified as shrubland or herbaceous grassland in NLCD were assigned new classes based on the Landfire Existing Vegetation Type (EVT) classification system. In total, the LUCAS STSM contained 61 unique LULC classes (Fig 1). 
We modeled transitions between state classes to represent major LULC change processes, including urbanization, agricultural expansion and contraction, forest management (i.e. clear-cut and selection harvest), wildfire, and insect damage. For wildfire and insect damage, we partitioned transition events into low, medium and high severity events. For land use transitions (urbanization, agricultural expansion and contraction), we generated spatial maps of the location of changes based on the NLCD time-series maps for the years 2001, 2006, 2011, and 2016. The 5-year change rates were then annualized, and the model was parameterized to stochastically select cells to transition from one class to another within the 5-year period. For the forest management and disturbance transitions, we used the annual time-series maps from the Landfire Disturbance database to identify areas of change. For disturbance transitions, we assumed that all cells would transition back to their original LULC class in the year following the disturbance event. Forest mortality resulting from drought and/or insects were derived from U.S. Forest Service Aerial Detection Surveys and were modeled the same was as fire and harvest transitions. 

We adopted the carbon stock and flux structure of the CBM-CFS3 model for this study. The approach includes the use of five live carbon pools (including above- and belowground pools) and 9 dead organic matter (DOM) pools covering standing dead trees, down deadwood, litter and carbon stored in soils. DOM pools are organized and named based on their rate of decay (e.g. very fast, fast, medium, slow). Carbon transfers between pools followed the same convention as the CBM-CFS3 model with two important modifications. First, rather than using net growth increment, we calculated net primary production (NPP) for each forest type-group and age and used this to drive annual gross carbon accumulation. Second, we introduced annual spatial multipliers to scale NPP based on variations in local climate conditions. These modifications are discussed in more detail below. 




### CBM-CFS3 reference simulations
We used the CBM-CFS3 model (version 1.2) to generate estimates of carbon stocks across all live and DOM pools for a 1-ha representative stand for each forest type-group. Each forest type-group was assigned to the closest forest type found in the CBM database; when no direct type was available we used generic hardwood or softwood types from CBM-CFS3. Additionally, each species was assigned to a representative ecozone and administrative boundary. We assumed wildfire was the historic stand replacing disturbance type and was the most recent disturbance while using default historic return intervals from CBM-CFS3. No additional disturbances were modeled for the reference simulation. Lastly, we calculated the mean temperature across the range of each forest type-group, using the value to modify the CBM-CFS3 reference simulation. 

#### Merchantable volume curves
The CBM-CFS3 model relies on users to provide merchantable volume curves for each tree species modeled. For each forest type-group, parameters used to estimate merchantable volume were queried from the U.S. Forest Service Forest Inventory and Analysis (FIA) database. We used the Von Bertalanffy growth equation to estimate merchantable volume by age as:

$$ y = a * {1-} exp{(-b * age)}^3 $$

where *y* is the merchantable volume, *a* is the asymptote and *b* is the rate of approach to the asymptote. The resulting estimates of merchantable volume by age (0-300 years), along with default biomass expansion factors from the CBM-CFS3 database, were used as inputs for the reference simulations. The CBM-CFS3 model was then run on an annual timestep for 300 years to estimate the amount of carbon stored in each pool by age by applying species-specific expansion factors to estimate biomass in each tree component. A carbon fraction was applied to estimate the carbon portion of the biomass stock. We used the default biomass expansion factors and carbon proportions from the CBM-CFS3 database.


```{r crosswalk, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
crosswalk_eco = read.csv("data/tables/cbm-lucas-lookup.csv") 
datatable(crosswalk_eco, options = list(pageLength = 10, autoWidth = TRUE), rownames= FALSE)

# kable(crosswalk_eco, 
#       digits = 2, 
#       booktabs = T, 
#       escape = F, 
#       caption="Crosswalk table used to connect U.S. forest type-groups to Ecozones, Provinces and species types.") %>% 
#   kable_styling(latex_options = c("hold_position"), bootstrap_options = c("striped"))
```



### Carbon flow rates: LUCAS Flow Pathways module

We developed a sub-module within LUCAS to calculate carbon flux parameters based on output from the reference simulations and the CBM-CFS3 database. The module uses the species crosswalk table from above, along with crosswalk tables linking carbon stocks and disturbance types used in LUCAS and CBM-CFS3. Within the LCUAS model we specified a set of carbon Flow Pathways defining all of the carbon pools and fluxes consistent with those used in the CBM-CFS3. 


#### Net Primary Productivity
The module first uses results from the CBM-CFS3 reference simulation to calculate net primary productivity (NPP) as the sum of net growth and biomass turnover for a given age:

$$ NPP = \sum_{a} G_{(f,b,s,fr,cr)} + \sum_{a-1} T_{(f,b,s,fr,cr)}$$
where $G$ is net growth, $T$ is biomass turnover, $f$ is foliage, $b$ is branches and other wood, $s$ is merchantable stems, $fr$ if fine roots, $cr$ is coarse roots and $a$ is forest age. Thus, outputs consist of an estimate of NPP by age (up to 300 years old) and are stored as state attributes for each state class type (i.e. forest type-group). Results are stored as a State Attribute within the LUCAS model.

Output from the CBM-CFS3 reference simulation also allows us to calculate the proportional allocation of NPP, by age, between the five live tree component stocks.

$$ pNPP_{s_i}^{a} = \frac{G_{s_i}^{a} + (S_{s_i}^{{a-1}} * T_{s_i})} {NPP^a}  $$
where $pNPP_{s_i}^{a^t}$ is the proportion of NPP allocated to one of five live carbon pools ($s_i$) for a given aged forest ($a^t$); $G_{s_i}^{a^t}$ is the net growth of carbon pool $i$ at age $t$; $S_{s_i}^{a^{t-1}}$ is the amount of carbon stored in pool $i$ at age $t-1$, $T_{s_i}$ is the species and region specific carbon turnover rate, and $NPP^a$ is the total species and region specific NPP for each forest age. Results are stored as Flow Multipliers within the LUCAS model and used to allocate annual NPP State Attribute values to each live tree component.

#### Carbon flux rates
The LUCAS model uses a set of flow multipliers to estimate carbon fluxes representing biomass turnover, decay and decomposition, emissions, and transfers of carbon resulting from LULC change and disturbance. The LUCAS Flow Pathwyas module uses the set of crosswalk tables described above to parameterize a flow multipliers table for each unique combination of forest type-group and ecozone. Flow multipliers specify the annual rate and proportion of carbon transferred from one stock type to another. The ordering of carbon fluxes within the LUCAS model was established to match that of the CBM-CFS3.

1. Transfer of snag stems and branches to down deadwood pools (aboveground medium and fast, respectively),
2. Emission and decay of standing (snag stems and branches) and down deadwood (aboveground medium),
3. Emission from the belowground slow pool,
4. Biomass turnover from live to DOM pools
5. Emission and decay from DOM pools (belowground very fast, belowground fast, aboveground very fast, aboveground fast)
6. Emission from aboveground slow
7. Transfer from aboveground slow to belowground slow
8. Growth of live pools.

In addition to the base carbon flux rates, the LUCAS Flow Pathways module also parameterzies a set of transition-based flow multipliers which control the rate at which carbon is transferred between pools when a change in LULC or disturbance occurs. Transition triggered flows are implemented at the end of each timestep after all base flows have occured. For this study, we considered the effects of fire (high, medium, and low severity), harvest (clearcut and selection), drought/insect mortality, urbanization, agricultural expansion (i.e. deforestation), and agricultural contraction (i.e. reforestation). Flow rates were imported from transition matricies from the CBM-CFS3 model.








### Model validation
Within the LUCAS model we ran a 300 year simulation with all cells starting at age 0 using the carbon flux rates to estimate changes in stocks developed in the previous step. A single 1-ha representative stand was run for each forest group-type. No disturbances were simulated. We compared individual carbon stock output from the LUCAS simulation with the original output from the CBM-CFS3 runs to ensure we could reliably reproduce results.

### Spin-up of Dead Organic Matter (DOM) Pools
We used the LUCAS model to spin-up DOM pools over a 3000 year simulation period under two assumptions. Fire was assumed as the historical stand replacing disturbance in both scenarios with a replacement interval specified for each forest type-group and ecozone. We then imposed both wildfire and clearcut harvest as the most recent disturbance and ran the model for another 300 years in order to generate carbon over age look-up tables for each assumption (harvest and clearcut as last disturbances). The final 300 years worth of data were used to initialize carbon for each land cell in the conterminous U.S..

### Mapping Initial Carbon Stocks in 2001
Results from the LUCAS spin-up were used to initialize carbon stocks for each 1-km x 1-km cell in the conterminous U.S. for the year 2001. State class maps (LULC + forest type-group) and forest age were developed and each cell was initialized with carbon stocks for both fire and clearcut harvest as the last disturbance. We used the National Interagency Fire Consortium's map of historical fire perimeters to identify cells which should have their final initialization based on fire; if a cell was not previously burned it was initiated with stocks from the clearcut harvest simulation.




































